---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Before starting we have to load some libraries neccessary for next work. Also, we set a name of directory, in which our dataset is.
```{r settingThingsUp, echo=TRUE, results='hide'}
library(dplyr)
library(lattice)
filename <- "activity.zip"
```


## Loading and preprocessing the data
```{r loadAndPreprocess, echo=TRUE}

data <- read.csv(unz(filename, "activity.csv"),
                 header=TRUE,
                 colClasses=c("numeric","Date","numeric"))

data <- tbl_df(data)

with.NAs <- data
without.NAs <- filter(data,
                      !is.na(steps))
```


## What is mean total number of steps taken per day?
```{r meanMedianAndHistogram, echo=TRUE}

# we must group data by date
data.1 <- group_by(without.NAs,
                   date)

# the next step is adding new variable, which contains sum of steps by days
data.1 <- mutate(data.1,
                 steps.by.days=sum(steps))

# we want unique days, because each day with one date has the same sum of steps  
data.1 <- distinct(data.1,
                   date)

# finally, we can draw a histogram
hist(data.1$steps.by.days,
     xlab="Number of Steps Taken",
     ylab="Number of Days",
     main="Histogram of number of steps taken by days",
     labels=TRUE,
     col="orange")

data1.mean <- mean(data.1$steps.by.days)
data1.median <- median(data.1$steps.by.days)
```
Mean total number of steps taken per day is `r data1.mean`, while the median has value of `r data1.median`.


## What is the average daily activity pattern?
```{r averageDailyPattern, echo=TRUE}

# we must group data by intervals
data.2 <- group_by(without.NAs,
                   interval)

# the next step is adding new variable, which contains average of steps by intervals
data.2 <- mutate(data.2,
                 average.steps.by.intervals=sum(steps)/n())

# we want unique inervals, because each interval with the same number has the same average of steps  
data.2 <- distinct(data.2,
                   interval)

# finally, we can draw a plot
with(data.2, plot(interval,
                  average.steps.by.intervals,
                  type="l",
                  xlab="Intervals",
                  ylab="Average Number of Steps Taken"))

# we must ungroup the data so that we can find interval with maximal average steps taken
data.2 <- ungroup(data.2)

# now we can find needed interval
maximum.number.of.steps <- filter(data.2,
                                  average.steps.by.intervals==max(average.steps.by.intervals))$interval
```
Interval **`r maximum.number.of.steps`** contains the maximum number of steps, on average across all the days in the dataset.


## Imputing missing values
```{r imputingMissingValues, echo=TRUE}
sum.of.NAs <- sum(is.na(with.NAs))
```
The total number of missing values in the dataset (i.e. the total number of rows with `NA`s) is **`r sum.of.NAs`**.

In the next code chunk we firstly group data by inervals (including rows with `NA`s). That is made because we are going to replace all `NA`s with mean number of steps taken in each interval.
```{r}
data.3 <- group_by(with.NAs, interval)
```

So after grouping the data is done, we add a variable called inerval.mean, which contains mean number of steps taken for each interval separately.
```{r}
data.3 <- mutate(data.3, interval.mean=mean(steps, na.rm=TRUE))
```

The last step is done with ease of `ifelse` statement, which allows us to simply replace `NA`s with mean number of steps for current interval, or to let there original number of steps.
```{r}
data.3 <- mutate(data.3, steps=ifelse(is.na(steps), interval.mean, steps))
data.3 <- select(data.3, steps, date, interval)
```


```{r}
data.3 <- ungroup(data.3)

data.3 <- group_by(data.3,
                   date)

data.3 <- mutate(data.3,
                 steps.by.days=sum(new.steps))

data.3 <- distinct(data.3,
                   date)

hist(data.3$steps.by.days,
     xlab="Number of Steps Taken",
     ylab="Number of Days",
     main="Histogram of number of steps taken by days",
     labels=TRUE,
     col="orange")

data3.mean <- mean(data.3$steps.by.days)
data3.median <- median(data.3$steps.by.days)

```
Mean total number of steps taken per day with `NA`s replaced is `r data3.mean`, while the median has value of `r data3.median`.


## Are there differences in activity patterns between weekdays and weekends?
In this part we have got one restriction, which is to use the dataset made few steps before, where all `NA`s were replaced.
```{r}

# we simply add a variable containing type of the current day
data.4 <- mutate(without.NAs,
                 day.type=ifelse(weekdays(date) %in% c('sobota','nedela'),'weekend','weekday'))

# we must group data by intervals and by date.type simultaneously
data.4 <- group_by(data.4,
                   interval,
                   day.type)

data.4 <- mutate(data.4,
                 average.steps.by.day.type=sum(steps)/n())


data.4 <- distinct(data.4,
                   interval,
                   day.type)


xyplot(average.steps.by.day.type ~ interval | day.type,
       data=data.4,
       xlab="Intervals",
       ylab="Average Number of Steps Taken",
       layout=c(1,2),
       type="l")

data.4 <- ungroup(data.4)

filter(data.4,
       average.steps.by.day.type==max(average.steps.by.day.type))
```